---
title: "R Notebook"
output: html_notebook
---


## Import libraries
```{r}
library(tidymodels)
```

## Import dataset
```{r}
df.train <- Train
df.test <- Test
head(df.train)
```

```{r}
tail(df.train)
```

```{r}
glimpse(df.train)
```

```{r}
skimr::skim(df.train)
```


```{r}
is.na(df.train) %>% colSums()
```


```{r}
df.train <- df.train %>% 
            mutate(Label = factor(Label))
glimpse(df.train)
```

```{r}
names(df.train)
```
## Modeling


## Recipe
```{r}
df_rec <-
  recipe(Label ~.,
         data = df.train) %>%
   update_role(ID, new_role = "ID") %>%
  step_corr(all_predictors(), threshold = 0.7, method = "spearman") 
summary(df_rec)
```

##
```{r}
prepped_data <- 
  df_rec %>% # use the recipe object
  prep() %>% # perform the recipe on training data
  juice() # extract only the preprocessed dataframe 
```

##
```{r}
glimpse(prepped_data)
```

## Folds
```{r}
set.seed(1001)

cv_folds <-
 vfold_cv(df.train, 
          v = 5, 
          strata = Label ) 
```



## Modeling
#4.*Logistic regression*

#Specify model
```{r}
log_spec <- # your model specification
  logistic_reg() %>%  # model type
  set_engine(engine = "glm") %>%  # model engine
  set_mode("classification") # model mode

# Show your model specification
log_spec
```

#Creating Workflow
```{r}
log_wflow <- # new workflow object
 workflow() %>% # use workflow function
 add_recipe(df_rec) %>%   # use the new recipe
 add_model(log_spec)   # add your model spec

# show object
log_wflow
```
```{r}
log_res <- 
  log_wflow %>% 
  fit_resamples(
    resamples = cv_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(
      save_pred = TRUE)
    ) 
```







#
```{r}
# save model coefficients for a fitted model object from a workflow

get_model <- function(x) {
  pull_workflow_fit(x) %>% tidy()
}

# same as before with one exception
log_res_2 <- 
  log_wflow %>% 
  fit_resamples(
    resamples = cv_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(
      save_pred = TRUE,
      extract = get_model) # use extract and our new function
    ) 
```


#Evaluate model
```{r}
all_coef <- map_dfr(log_res_2$.extracts, ~ .x[[1]][[1]])
```

#
```{r}
filter(all_coef, term == "DEROG")
```

#Performance metrics
##Show average performance over all folds
```{r}
log_res %>%  collect_metrics(summarize = TRUE)
```

##Show performance for every single fold:


```{r}
log_res %>%  collect_metrics(summarize = FALSE)

```

#Collect predictions

```{r}
#our prediction saved as log_pred
log_pred <- 
  log_res %>%
  #function to collect obtain the actual model predictions
  collect_predictions()
```

##Confusion matrix
```{r}
log_pred %>% 
  conf_mat(Label, .pred_class) 
```
##A quick visualizion of  our confusion matrix
```{r}
log_pred %>% 
  conf_mat(Label, .pred_class) %>% 
  autoplot(type = "mosaic")
```

## Heatmap of our confusion matrix
```{r}
log_pred %>% 
  conf_mat(Label, .pred_class) %>% 
  autoplot(type = "heatmap")
```
```{r}
head(log_pred)
```

##ROC-Curve
ROC curve for our 5 folds
```{r}
log_pred %>% 
  group_by(id) %>% # id contains our folds
  roc_curve(Label, .pred_0) %>% 
  autoplot()
```

#Probability distributions
Plot predicted probability distributions for our two classes.
```{r}
log_pred %>% 
  ggplot() +
  geom_density(aes(x = .pred_0, 
                   fill = Label), 
               alpha = 0.5)
```
#5. Random forest

## Specify models
```{r}
library(ranger)

rf_spec <- 
  rand_forest() %>% 
  set_engine("ranger", importance = "impurity") %>% 
  set_mode("classification")
```


## Bundle recipe and model:
```{r}
rf_wflow <-
 workflow() %>%
 add_recipe(df_rec) %>% 
 add_model(rf_spec) 
```

## performance metrics.
```{r}
rf_res <-
  rf_wflow %>% 
  fit_resamples(
    resamples = cv_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

rf_res %>%  collect_metrics(summarize = TRUE)
```


## Model coefficients
```{r}
# save model coefficients for a fitted model object from a workflow

get_model <- function(x) {
  pull_workflow_fit(x) %>% tidy()
}

# same as before with one exception
rf_res_2 <- 
  rf_wflow %>% 
  fit_resamples(
    resamples = cv_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(
      save_pred = TRUE,
      extract = get_model) # use extract and our new function
    )
```

## Performance metrics


### Show average performance over all folds
```{r}
rf_res %>%  collect_metrics(summarize = TRUE)
```


### Show performance for every single fold:
```{r}
rf_res %>%  collect_metrics(summarize = FALSE)
```

#### Collect predictions

```{r}
rf_pred <- 
  rf_res %>%
  collect_predictions()
```

#### Confusion matrix
```{r}
rf_pred %>% 
  conf_mat(Label, .pred_class) 
```

###Quick visualization confusion matrix
```{r}
rf_pred %>% 
  conf_mat(Label, .pred_class) %>% 
  autoplot(type = "mosaic")
```
####Confusion matrix Heatmap
```{r}
rf_pred %>% 
  conf_mat(Label, .pred_class) %>% 
  autoplot(type = "heatmap")
```
###ROC-Curve
ROC curve for our 5 folds
```{r}
rf_pred %>% 
  group_by(id) %>% # id contains our folds
  roc_curve(Label, .pred_0) %>% 
  autoplot()
```

###Probability distributions
```{r}
rf_pred %>% 
  ggplot() +
  geom_density(aes(x = .pred_0, 
                   fill = Label), 
               alpha = 0.5)
```

#6:XGBoost
##Boosted tree (XGBoost)
#specify model
```{r}
library(xgboost)

xgb_spec <- 
  boost_tree() %>% 
  set_engine("xgboost") %>% 
  set_mode("classification") 
```
#Bundle recipe and model with workflows:
```{r}
xgb_wflow <-
 workflow() %>%
 add_recipe(df_rec) %>% 
 add_model(xgb_spec)
```




#
```{r}
xgb_res <- 
  xgb_wflow %>% 
  fit_resamples(
    resamples = cv_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(
      save_pred = TRUE)
    ) 
```


##metrics
```{r}
xgb_res <- 
  xgb_wflow %>% 
  fit_resamples(
    resamples = cv_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

xgb_res %>% collect_metrics(summarize = TRUE)
```


#7:K-nearest neighbor

# Specify model
```{r}
knn_spec <- 
  nearest_neighbor(neighbors = 4) %>% # we can adjust the number of neighbors 
  set_engine("kknn") %>% 
  set_mode("classification") 
```

#Bundle recipe and model with workflows:

```{r}
knn_wflow <-
 workflow() %>%
 add_recipe(df_rec) %>% 
 add_model(knn_spec)
```

#
```{r}
knn_res <- 
  knn_wflow %>% 
  fit_resamples(
    resamples = cv_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(
      save_pred = TRUE)
    ) 
```

#
```{r}
knn_res <- 
  knn_wflow %>% 
  fit_resamples(
    resamples = cv_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

knn_res %>% collect_metrics(summarize = TRUE)
```


ROC curve for our 5 folds
```{r}
rf_pred <- 
  knn_res %>%
  collect_predictions()

rf_pred %>% 
  group_by(id) %>% # id contains our folds
  roc_curve(Label, .pred_0) %>% 
  autoplot
```
#9.Compare models
Extract metrics from our models to compare them:


```{r}
library(forcats) #for fct_reorder function
log_metrics <- 
  log_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "Logistic Regression") # add the name of the model to every row

rf_metrics <- 
  rf_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "Random Forest")

xgb_metrics <- 
  xgb_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "XGBoost")

knn_metrics <- 
  knn_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "Knn")

# nnet_metrics <- 
#   nnet_res %>% 
#   collect_metrics(summarise = TRUE) %>%
#   mutate(model = "Neural Net")

# create dataframe with all models
model_compare <- bind_rows(
                          log_metrics,
                           rf_metrics,
                           xgb_metrics,
                           knn_metrics,
                         # nnet_metrics
                           ) 

# change data structure
model_comp <- 
  model_compare %>% 
  select(model, .metric, mean, std_err) %>% 
  pivot_wider(names_from = .metric, values_from = c(mean, std_err)) 
 
# show mean F1-Score for every model
model_comp %>% 
  arrange(mean_f_meas) %>% 
  mutate(model = fct_reorder(model, mean_f_meas)) %>% # order results
  ggplot(aes(model, mean_f_meas, fill=model)) +
  geom_col() +
  coord_flip() +
  scale_fill_brewer(palette = "Blues") +
   geom_text(
     size = 3,
     aes(label = round(mean_f_meas, 2), y = mean_f_meas + 0.08),
     vjust = 1
  )

```
`OBSERVATION`:knn model performed best in our train set hence will be used on test set.

```{r}
# show mean area under the curve (auc) per model
model_comp %>% 
  arrange(mean_roc_auc) %>% 
  mutate(model = fct_reorder(model, mean_roc_auc)) %>%
  ggplot(aes(model, mean_roc_auc, fill=model)) +
  geom_col() +
  coord_flip() +
  scale_fill_brewer(palette = "Blues") + 
     geom_text(
     size = 3,
     aes(label = round(mean_roc_auc, 2), y = mean_roc_auc + 0.08),
     vjust = 1
  )
```
 Let’s find the maximum mean F1-Score:
```{r}
model_comp %>% slice_max(mean_f_meas)

```



## Making predictions
we Will use random forest
```{r}


final_wf <- finalize_workflow(rf_wflow, rf_spec)
final_mod <- final_wf %>% fit(df.train)



predict_res <- predict(
        final_mod,
        df.test)



```

###

```{r}
result_2 <- data.frame(df.test$ID,predict_res$.pred_class)
glimpse(result_2)


#write.csv(result, "C:/Users/OMBATI/Desktop/R codes/UmojaHack\\dkutR2.csv", row.names=FALSE)

```

```{r}
result_2 <- result_2 %>%
          rename(ID = df.test.ID  ,Target = predict_res..pred_class )
result_2
```

```{r}
write.csv(result_2, "C:/Users/OMBATI/Desktop/R codes/UmojaHack\\dkutR3.csv", row.names=FALSE)
```

